---
title: "Code"
author: "Yaxi Lou"
date: "June 8, 2017"
output: pdf_document
---

## xgboost
### preparing data set
\textbf{xgboost requres a numeric matrix, so we have to recode categorical variables to numeric variables based on the levels and frequencies.}  
```{r}
# read dataset
library(data.table)
lafd_train<-fread("lafdtraining.csv")
lafd_test<-fread("testing.without.response.txt",sep=",")
```

\textbf{We find test data contains NAs under variable Dispatch Sequence, replace it by 0.}  
```{r}
train<-lafd_train
test<-lafd_test
rm(lafd_train)
rm(lafd_test)
## We find test data contains NAs under variable Dispatch Sequence, replace it by 0
for(i in 1:nrow(test)){
if(is.na(test$`Dispatch Sequence`[i]))
  test$`Dispatch Sequence`[i]<-0
}
train[which(is.na(train$`Dispatch Sequence`)),6]<-0
## delete NAs from vatiable elapsed_time in train
train<-train[-which(is.na(train$elapsed_time)),]
## delete NAs from variable PPE Level in train
train<-train[-which(is.na(train$`PPE Level`)),]
```


\textbf{We find the corresponding zip code to each station.}  

```{r}
train$zip<-0
train[which(train$`First in District`==1),]$zip<-90031
train[which(train$`First in District`==2),]$zip<-90033
train[which(train$`First in District`==3),]$zip<-90012
train[which(train$`First in District`==4),]$zip<-90012
train[which(train$`First in District`==5),]$zip<-90045
train[which(train$`First in District`==6),]$zip<-90004
train[which(train$`First in District`==7),]$zip<-91331
train[which(train$`First in District`==8),]$zip<-91326
train[which(train$`First in District`==9),]$zip<-90014
train[which(train$`First in District`==10),]$zip<-90015
train[which(train$`First in District`==11),]$zip<-90057
train[which(train$`First in District`==12),]$zip<-90042
train[which(train$`First in District`==13),]$zip<-90006
train[which(train$`First in District`==14),]$zip<-90011
train[which(train$`First in District`==15),]$zip<-90007
train[which(train$`First in District`==16),]$zip<-90032
train[which(train$`First in District`==17),]$zip<-90021
train[which(train$`First in District`==18),]$zip<-91344
train[which(train$`First in District`==19),]$zip<-90049
train[which(train$`First in District`==20),]$zip<-90026
train[which(train$`First in District`==21),]$zip<-90011
train[which(train$`First in District`==23),]$zip<-90272
train[which(train$`First in District`==24),]$zip<-91040
train[which(train$`First in District`==25),]$zip<-90023
train[which(train$`First in District`==26),]$zip<-90018
train[which(train$`First in District`==27),]$zip<-90028
train[which(train$`First in District`==28),]$zip<-91326
train[which(train$`First in District`==29),]$zip<-90010
train[which(train$`First in District`==33),]$zip<-90003
train[which(train$`First in District`==34),]$zip<-90019
train[which(train$`First in District`==35),]$zip<-90027
train[which(train$`First in District`==36),]$zip<-90732
train[which(train$`First in District`==37),]$zip<-90024
train[which(train$`First in District`==38),]$zip<-90744
train[which(train$`First in District`==39),]$zip<-91401
train[which(train$`First in District`==40),]$zip<-90731
train[which(train$`First in District`==41),]$zip<-90046
train[which(train$`First in District`==42),]$zip<-90041
train[which(train$`First in District`==43),]$zip<-90034
train[which(train$`First in District`==44),]$zip<-90065
train[which(train$`First in District`==46),]$zip<-90037
train[which(train$`First in District`==47),]$zip<-90032
train[which(train$`First in District`==48),]$zip<-90731
train[which(train$`First in District`==49),]$zip<-90744
train[which(train$`First in District`==50),]$zip<-90065
train[which(train$`First in District`==51),]$zip<-90045
train[which(train$`First in District`==52),]$zip<-90029
train[which(train$`First in District`==55),]$zip<-90041
train[which(train$`First in District`==56),]$zip<-90039
train[which(train$`First in District`==57),]$zip<-90044
train[which(train$`First in District`==58),]$zip<-90035
train[which(train$`First in District`==59),]$zip<-90025
train[which(train$`First in District`==60),]$zip<-91601
train[which(train$`First in District`==61),]$zip<-90036
train[which(train$`First in District`==62),]$zip<-90066
train[which(train$`First in District`==63),]$zip<-90291
train[which(train$`First in District`==64),]$zip<-90061
train[which(train$`First in District`==65),]$zip<-90002
train[which(train$`First in District`==66),]$zip<-90047
train[which(train$`First in District`==67),]$zip<-90094
train[which(train$`First in District`==68),]$zip<-90019
train[which(train$`First in District`==69),]$zip<-90272
train[which(train$`First in District`==70),]$zip<-91324
train[which(train$`First in District`==71),]$zip<-90024
train[which(train$`First in District`==72),]$zip<-91303
train[which(train$`First in District`==73),]$zip<-91335
train[which(train$`First in District`==74),]$zip<-91042
train[which(train$`First in District`==75),]$zip<-91345
train[which(train$`First in District`==76),]$zip<-90068
train[which(train$`First in District`==77),]$zip<-91352
train[which(train$`First in District`==78),]$zip<-91604
train[which(train$`First in District`==79),]$zip<-90248
train[which(train$`First in District`==81),]$zip<-91402
train[which(train$`First in District`==82),]$zip<-90028
train[which(train$`First in District`==83),]$zip<-91436
train[which(train$`First in District`==84),]$zip<-91367
train[which(train$`First in District`==85),]$zip<-90710
train[which(train$`First in District`==86),]$zip<-91602
train[which(train$`First in District`==87),]$zip<-91343
train[which(train$`First in District`==88),]$zip<-91403
train[which(train$`First in District`==89),]$zip<-91605
train[which(train$`First in District`==90),]$zip<-91406
train[which(train$`First in District`==91),]$zip<-91342
train[which(train$`First in District`==92),]$zip<-90064
train[which(train$`First in District`==93),]$zip<-91356
train[which(train$`First in District`==94),]$zip<-90008
train[which(train$`First in District`==95),]$zip<-90045
train[which(train$`First in District`==96),]$zip<-91311
train[which(train$`First in District`==97),]$zip<-90046
train[which(train$`First in District`==98),]$zip<-91331
train[which(train$`First in District`==99),]$zip<-90210
train[which(train$`First in District`==100),]$zip<-91406
train[which(train$`First in District`==101),]$zip<-90732
train[which(train$`First in District`==102),]$zip<-91401
train[which(train$`First in District`==103),]$zip<-91325
train[which(train$`First in District`==104),]$zip<-91306
train[which(train$`First in District`==105),]$zip<-91367
train[which(train$`First in District`==106),]$zip<-91304
train[which(train$`First in District`==107),]$zip<-91311
train[which(train$`First in District`==108),]$zip<-90210
train[which(train$`First in District`==109),]$zip<-90049
train[which(train$`First in District`==112),]$zip<-90731
```

```{r}
test$zip<-0
test[which(test$`First in District`==1),]$zip<-90031
test[which(test$`First in District`==2),]$zip<-90033
test[which(test$`First in District`==3),]$zip<-90012
test[which(test$`First in District`==4),]$zip<-90012
test[which(test$`First in District`==5),]$zip<-90045
test[which(test$`First in District`==6),]$zip<-90004
test[which(test$`First in District`==7),]$zip<-91331
test[which(test$`First in District`==8),]$zip<-91326
test[which(test$`First in District`==9),]$zip<-90014
test[which(test$`First in District`==10),]$zip<-90015
test[which(test$`First in District`==11),]$zip<-90057
test[which(test$`First in District`==12),]$zip<-90042
test[which(test$`First in District`==13),]$zip<-90006
test[which(test$`First in District`==14),]$zip<-90011
test[which(test$`First in District`==15),]$zip<-90007
test[which(test$`First in District`==16),]$zip<-90032
test[which(test$`First in District`==17),]$zip<-90021
test[which(test$`First in District`==18),]$zip<-91344
test[which(test$`First in District`==19),]$zip<-90049
test[which(test$`First in District`==20),]$zip<-90026
test[which(test$`First in District`==21),]$zip<-90011
test[which(test$`First in District`==23),]$zip<-90272
test[which(test$`First in District`==24),]$zip<-91040
test[which(test$`First in District`==25),]$zip<-90023
test[which(test$`First in District`==26),]$zip<-90018
test[which(test$`First in District`==27),]$zip<-90028
test[which(test$`First in District`==28),]$zip<-91326
test[which(test$`First in District`==29),]$zip<-90010
test[which(test$`First in District`==33),]$zip<-90003
test[which(test$`First in District`==34),]$zip<-90019
test[which(test$`First in District`==35),]$zip<-90027
test[which(test$`First in District`==36),]$zip<-90732
test[which(test$`First in District`==37),]$zip<-90024
test[which(test$`First in District`==38),]$zip<-90744
test[which(test$`First in District`==39),]$zip<-91401
test[which(test$`First in District`==40),]$zip<-90731
test[which(test$`First in District`==41),]$zip<-90046
test[which(test$`First in District`==42),]$zip<-90041
test[which(test$`First in District`==43),]$zip<-90034
test[which(test$`First in District`==44),]$zip<-90065
test[which(test$`First in District`==46),]$zip<-90037
test[which(test$`First in District`==47),]$zip<-90032
test[which(test$`First in District`==48),]$zip<-90731
test[which(test$`First in District`==49),]$zip<-90744
test[which(test$`First in District`==50),]$zip<-90065
test[which(test$`First in District`==51),]$zip<-90045
test[which(test$`First in District`==52),]$zip<-90029
test[which(test$`First in District`==55),]$zip<-90041
test[which(test$`First in District`==56),]$zip<-90039
test[which(test$`First in District`==57),]$zip<-90044
test[which(test$`First in District`==58),]$zip<-90035
test[which(test$`First in District`==59),]$zip<-90025
test[which(test$`First in District`==60),]$zip<-91601
test[which(test$`First in District`==61),]$zip<-90036
test[which(test$`First in District`==62),]$zip<-90066
test[which(test$`First in District`==63),]$zip<-90291
test[which(test$`First in District`==64),]$zip<-90061
test[which(test$`First in District`==65),]$zip<-90002
test[which(test$`First in District`==66),]$zip<-90047
test[which(test$`First in District`==67),]$zip<-90094
test[which(test$`First in District`==68),]$zip<-90019
test[which(test$`First in District`==69),]$zip<-90272
test[which(test$`First in District`==70),]$zip<-91324
test[which(test$`First in District`==71),]$zip<-90024
test[which(test$`First in District`==72),]$zip<-91303
test[which(test$`First in District`==73),]$zip<-91335
test[which(test$`First in District`==74),]$zip<-91042
test[which(test$`First in District`==75),]$zip<-91345
test[which(test$`First in District`==76),]$zip<-90068
test[which(test$`First in District`==77),]$zip<-91352
test[which(test$`First in District`==78),]$zip<-91604
test[which(test$`First in District`==79),]$zip<-90248
test[which(test$`First in District`==81),]$zip<-91402
test[which(test$`First in District`==82),]$zip<-90028
test[which(test$`First in District`==83),]$zip<-91436
test[which(test$`First in District`==84),]$zip<-91367
test[which(test$`First in District`==85),]$zip<-90710
test[which(test$`First in District`==86),]$zip<-91602
test[which(test$`First in District`==87),]$zip<-91343
test[which(test$`First in District`==88),]$zip<-91403
test[which(test$`First in District`==89),]$zip<-91605
test[which(test$`First in District`==90),]$zip<-91406
test[which(test$`First in District`==91),]$zip<-91342
test[which(test$`First in District`==92),]$zip<-90064
test[which(test$`First in District`==93),]$zip<-91356
test[which(test$`First in District`==94),]$zip<-90008
test[which(test$`First in District`==95),]$zip<-90045
test[which(test$`First in District`==96),]$zip<-91311
test[which(test$`First in District`==97),]$zip<-90046
test[which(test$`First in District`==98),]$zip<-91331
test[which(test$`First in District`==99),]$zip<-90210
test[which(test$`First in District`==100),]$zip<-91406
test[which(test$`First in District`==101),]$zip<-90732
test[which(test$`First in District`==102),]$zip<-91401
test[which(test$`First in District`==103),]$zip<-91325
test[which(test$`First in District`==104),]$zip<-91306
test[which(test$`First in District`==105),]$zip<-91367
test[which(test$`First in District`==106),]$zip<-91304
test[which(test$`First in District`==107),]$zip<-91311
test[which(test$`First in District`==108),]$zip<-90210
test[which(test$`First in District`==109),]$zip<-90049
test[which(test$`First in District`==112),]$zip<-90731
```

\textbf{Recode other vatiables according to levels.}  
```{r}
table(train$`Emergency Dispatch Code`) #Emergency
table(train$`Dispatch Status`) #12 chr
table(train$`Unit Type`) #40 chr
table(train$`PPE Level`) # EMS Non-EMS
```

\textbf{Emergency Dispatch Code}  
```{r}
train$`Emergency Dispatch Code`<-1
train$`Emergency Dispatch Code`<-as.numeric(train$`Emergency Dispatch Code`)
test$Emergency.Dispatch.Code<-1
test$Emergency.Dispatch.Code<-as.numeric(test$Emergency.Dispatch.Code)
```

\textbf{Dispatch Status}  
```{r}
train$`Dispatch Status`<-as.factor(train$`Dispatch Status`)
num.class <- length(levels(train$`Dispatch Status`))
levels(train$`Dispatch Status`) <- 1:num.class
train$`Dispatch Status`<-as.numeric(train$`Dispatch Status`)

test$`Dispatch Status`<-as.factor(test$`Dispatch Status`)
num.class <- length(levels(test$`Dispatch Status`))
levels(test$`Dispatch Status`) <- 1:num.class
test$`Dispatch Status`<-as.numeric(test$`Dispatch Status`)
```

\textbf{Unit Type}  
```{r}
train$`Unit Type`<-as.factor(train$`Unit Type`)
num.class <- length(levels(train$`Unit Type`))
levels(train$`Unit Type`) <- 1:num.class
train$`Unit Type`<-as.numeric(train$`Unit Type`)

test$`Unit Type`<-as.factor(test$`Unit Type`)
num.class <- length(levels(test$`Unit Type`))
levels(test$`Unit Type`) <- 1:num.class
test$`Unit Type`<-as.numeric(test$`Unit Type`)
```

\textbf{PPE Level}  
```{r}
train$`PPE Level`<-as.factor(train$`PPE Level`)
num.class <- length(levels(train$`PPE Level`))
levels(train$`PPE Level`) <- 1:num.class
train$`PPE Level`<-as.numeric(train$`PPE Level`)

test$`PPE Level`<-as.factor(test$`PPE Level`)
num.class <- length(levels(test$`PPE Level`))
levels(test$`PPE Level`) <- 1:num.class
test$`PPE Level`<-as.numeric(test$`PPE Level`)
```

\textbf{convert int to numeric}  
```{r}
train$year<-as.numeric(train$year)
train$`First in District`<-as.numeric(train$`First in District`)
train$`Dispatch Sequence`<-as.numeric(train$`Dispatch Sequence`)

test$year<-as.numeric(test$year)
test$`First in District`<-as.numeric(test$`First in District`)
test$`Dispatch Sequence`<-as.numeric(test$`Dispatch Sequence`)
```

\textbf{Incident Creation Time (GMT)}  
```{r}
library(lubridate)

train$`Incident Creation Time (GMT)`<- hms(train$`Incident Creation Time (GMT)`)
train$hour <- hour(train$`Incident Creation Time (GMT)`)
train$minute <- minute(train$`Incident Creation Time (GMT)`)
train$hour <- ifelse(train$minute >= 40, train$hour + 1, train$hour)
train$hour <- ifelse(train$hour == 24, 0,train$hour)
train$timelevel <- ifelse(train$hour >=7 & train$hour <= 10, 1,
                         ifelse(train$hour >=11 & train$hour <= 14, 2,
                                ifelse(train$hour >=15 & train$hour <= 19, 3,
                                       ifelse(train$hour >=20 & train$hour <= 23, 4,5))))


test$`Incident Creation Time (GMT)`<- hms(test$`Incident Creation Time (GMT)`)
test$hour <- hour(test$`Incident Creation Time (GMT)`)
test$minute <- minute(test$`Incident Creation Time (GMT)`)
test$hour <- ifelse(test$minute >= 40, test$hour + 1, test$hour)
test$hour <- ifelse(test$hour == 24, 0,test$hour)
test$timelevel <- ifelse(test$hour >=7 & test$hour <= 10, 1,
                         ifelse(test$hour >=11 & test$hour <= 14, 2,
                                ifelse(test$hour >=15 & test$hour <= 19, 3,
                                       ifelse(test$hour >=20 & test$hour <= 23, 4,5))))
```

\textbf{Incident ID}  
```{r}
library(dplyr)
id_train<-train %>% group_by(incident.ID) %>% count()
train<-merge(train,id_train,by="incident.ID")

id_test<-test %>% group_by(incident.ID) %>% count()
test<-merge(test,id_test,by="incident.ID")
```

\textbf{traffic (There are heavier traffic during rush hour)}  
```{r}
train$traffic<-0
train[which(train$timelevel==1),]$traffic<-3
train[which(train$timelevel==2),]$traffic<-2
train[which(train$timelevel==3),]$traffic<-3
train[which(train$timelevel==4),]$traffic<-2
train[which(train$timelevel==5),]$traffic<-1

test$traffic<-0
test[which(test$timelevel==1),]$traffic<-3
test[which(test$timelevel==2),]$traffic<-2
test[which(test$timelevel==3),]$traffic<-3
test[which(test$timelevel==4),]$traffic<-2
test[which(test$timelevel==5),]$traffic<-1
```

### Apply xgboost model
\textbf{separate training and testing data}  
```{r}
set.seed(123)
i.train <- sample(nrow(train), nrow(train)*0.75,replace=FALSE)
Training_Inner<-train[i.train,]
Test_Inner<- train[-i.train,]
```

\textbf{preparing training set}  
```{r}
library(xgboost)
new_tr <- model.matrix(~.+0,data = Training_Inner[,c(3,6,9,12,16,17)]) 
new_ts <- model.matrix(~.+0,data = Test_Inner[,c(3,6,9,12,16,17)])
dtrain <- xgb.DMatrix(data = new_tr,label = Training_Inner$elapsed_time) 
dtest <- xgb.DMatrix(data = new_ts,label = Test_Inner$elapsed_time)
```

\textbf{model}  
```{r}
params <- list(booster = "gbtree", objective = "reg:linear", eta=0.3, gamma=0, max_depth=10, min_child_weight=1, subsample=1, colsample_bytree=1)

xgb1 <- xgb.train (params = params, data= dtrain, nrounds = 100, watchlist = list(train=dtrain), print_every_n = 10, early_stop_round = 10, maximize = F , eval_metric = "rmse")

xgbpred1<-predict(xgb1,new_ts)
mean((xgbpred1-Test_Inner$elapsed_time)^2)
```

\textbf{features importance}  
```{r}
library(ggplot2)
library(Ckmeans.1d.dp)
try<-xgb.importance(names(test[,-c(1,2)]),model=xgb1)
print(try)
xgb.ggplot.importance(try, measure = "Frequency", rel_to_first = TRUE)
```

\textbf{prediction}  
```{r} 
new_test <- model.matrix(~.+0,data = test[,c(6,8,9,11,12,17)])
xgbpred <- predict (xgb1,new_test)
prediction<-cbind(as.integer(test$row.id),xgbpred)
write.csv(prediction,file="prediction.csv")
```

## Random Forest
```{r}
lafd <- read_csv("lafdtraining.csv")
load("lafd_test.RData")
```

```{r}
list1 <- lafd[, c(1,7,8)]
trial1 <- merge(train, list1, by = "row.id")
list2 <- lafd_test[, c(1,7,8)]
test1 <- merge(test, list2, by = "row.id")
save(trial1, file = "mytrain1.RData")
save(test1, file = "mytest1.RData")
```

```{r}
trial2 <- trial1[, -c(2:5,7,8,10,13,14)]
test2 <- test1[, -c(2:5,7,8,10,12,13)]
names(trial2) <- c("row.id","sequence", "ppe", "elapsed_time", "zip", "timelevel", "n", "traffic","status","unit")
names(test2) <- c("row.id","sequence","ppe", "zip", "timelevel", "n", "traffic", "status","unit")
```

```{r}
list1 <- which(trial2$unit == "FC - Fire Chief"| trial2$unit == "HC - Hose Carrier" | trial2$unit == "PT - Planning Trailer")
trial2$unit[list1] <- "Other"
test2$unit <- as.character(test2$unit)
test2$unit[which(test2$unit == "RP - Rehab Plug Buggy")] <- "Other"
test2$unit <- as.factor(test2$unit)
```

```{r}
trial2$status <- ifelse(trial2$status == "OVD" | trial2$status =="CAV" | trial2$status == "WRS" | trial2$status =="PGR" | trial2$status =="TSP", "OTHER", trial2$status)
test2$status <- as.character(test2$status)
test2$status <- ifelse(test2$status == "OVD" | test2$status =="CAV" | test2$status == "WRS" | test2$status =="PGR" | test2$status =="TSP", "OTHER", test2$status)
trial2$unit <- as.factor(trial2$unit)
```

```{r}
trial2$speed <- ifelse(trial2$timelevel == 5 & trial2 $ traffic == 1, "low", 
                       ifelse(trial2$timelevel == 2 & trial2$traffic == 2, "medium", 
                              ifelse(trial2$timelevel == 4 & trial2$traffic == 2, "medium","high")))

test2$speed <- ifelse(test2$timelevel == 5 & test2 $traffic == 1, "low", 
                       ifelse(test2$timelevel == 2 &test2$traffic == 2, "medium", 
ifelse(test2$timelevel == 4 & test2$traffic == 2, "medium","high")))
```

```{r}
trial2 <- trial2[, -c(1)]
test2 <- test2[, -c(1,)]
```

\textbf{Apply model}
```{r}
library(h2o)
localH2O <- h2o.init(nthreads = -1)
h2o.removeAll()
```

```{r}
train.h2o <- as.h2o(trial3)
test.h2o <- as.h2o(test3)
colnames(train.h2o)
y.dep <- 3
x.indep <- c(1,2,4,5,6,7,8,9)
rforest.model <- h2o.randomForest(y=y.dep, x=x.indep, training_frame = train.h2o, ntrees = 500, mtries = 4, max_depth = 20, seed = 1000)
```

```{r}
h2o.performance(rforest.model)
h2o.varimp(rforest.model)
predict.rforest <- as.data.frame(h2o.predict(rforest.model, test.h2o))
```

